---
title: "Functions"
author: "Sonja"
date: "29/9/2021"
output: html_document
---

Functions accept input > perform taks, print, do a task > then when finished we might want them to return of value. 

```{r}

my_fun = function(x,y, print=TRUE)
{
  if(print)print("hello!")
  return(x>y)
}
  
```


Argument and Scope (what the function sees)

Inside the function body your function can see, everything in the global environment, values of any arguments, any temporary variables you create within the function body. 

Work on your function code outside of the function. when your code accomplishes what you want you can package it into the function
Use distinctive arguments. should not reference global variables. 

```{r}
require(palmerpenguins)
```

Build a function 
```{r}

image_file= "ugly_histogram.png"

png(
  here("images", image_file), 
  width = 1200, height = 1000
)

hist(penguins$flipper_length_mm)

dev.off()

```
make the function to save the png function 

```{r}
save_png_1=function(image_file)
{ 
  png(
  here("images", image_file), 
  width = 1200, height = 1000
  )
}

```


```{r}
save_png_1("ugly_histo_2.png")
hist(penguins$body_mass_g)

dev.off()

dat_vec= penguins$body_mass_g
my_title= "my histo"
x_label= "my data"

hist(
  dat_vec, 
  col="steelblue",
  main="sonj_title",
  xlab= x_label)

steelblue_fun=function(dat_vec, my_title, x_label)
{
  hist(
  dat_vec, 
  col="steelblue",
  main="sonj_title",
  xlab= x_label)
}

steelblue_fun(
  dat_vec = sample(x=1:100, size=100, replace=TRUE), 
  my_title = "sonj rand numbers", 
  x_label = "x-values"
)

```


