---
title: 'Lab 10: ANOVA by hand'
author: "Sonja"
date: "17/11/2021"
output: html_document
---

## Lab 10: ANOVA (by hand)

#### Worked with : 

Aliza Fassler
Justin Roch 

### Building an ANOVA by Hand

In this exercise you’ll get a chance to practice your R-coding skills and build an understanding of the inner workings of one-way Analysis of Variance by using R to do an ANOVA “by hand”!


#### Q1

```{r}
require(here)

rope = read.csv(here("data", "rope.csv"))
rope$rope.type = factor(rope$rope.type)

n_obs = nrow(rope)
n_groups = nlevels(rope$rope.type)

ss_tot = sum(((rope$p.cut-mean(rope$p.cut))^2))
df_tot = (nrow(rope)-1)

agg_resids = aggregate(
  x = rope$p.cut,
  by = list(rope$rope.type),
  FUN = function(x)x-mean(x))

agg_sq_resids = aggregate(x = rope$p.cut,
  by = list(rope$rope.type),
  FUN = function(x) sum((x - mean(x))^2))

ss_within = sum(agg_sq_resids$x)
df_within = n_obs-n_groups 


ss_among = ss_tot-ss_within
df_among = df_tot-df_within

ms_within = ss_within / (n_obs - n_groups)
ms_among  = ss_among / (n_groups - 1)
  

f_ratio = sum(ms_among/ms_within)
f_pval = 1 - pf(f_ratio, df_among, df_within)

```




#### Q1 self check 

```{r}
# number comparison tolerance
digits_check = 5

# Build the reference model using R functions
fit_1 = lm(p.cut ~ rope.type, data=rope)
anova(fit_1)
anova_fit_1 = anova(fit_1)

# Check degrees of freedom
anova_fit_1$Df == c(df_among, df_within)

# Check sums of squares
round(anova_fit_1$`Sum Sq`, digits = digits_check) == round(c(ss_among, ss_within), digits = digits_check)

# Check mean squares
round(anova_fit_1$`Mean Sq`, digits = digits_check) == round(c(ms_among, ms_within), digits = digits_check)

# Check the F-ratio
round(anova_fit_1$`F value`[1], digits = digits_check) == round(f_ratio, digits = digits_check)

# Check the F test statistic p-value
round(anova_fit_1$`Pr(>F)`[1], digits = digits_check) == round(f_pval, digits = digits_check)
```


#### Q2

Based on the figure I do not think there is equal variances in the rope type groups. There seems to be much more variance in the rope types BLAZE, SB, and VEL than the others. 



#### Q3

p value = 0.00143

```{r}
bartlett.test(p.cut ~ rope.type, data=rope)
```


#### Q4

By both visually examining the graph of the conditional boxplot and by conducting the Bartlett test, an ANOVA-type analysis does not seem appropriate for the raw data of the rope dataset. Having heterogeneous variance in the data violates the assumption of homogeneous variance between groups. 


```{r}
fit_rope_1 = lm(p.cut ~ rope.type, data = rope)
summary(fit_rope_1)
```



#### Q5 

The base case is rope type BLAZE. 



#### Q6

The mean percent cut is 0.36714, no calculation needed because it is simply the estimate of the coefficient of the base case. 



#### Q7 

```{r}
 0.36714 - 0.10164  
```
The percent cut rope type XTC is 0.2655. 
