---
title: "Lab 02: R Fundamentals 2"
author: "Sonja Glasser"
date: "15/9/2021"
output: html_document
---

Worked together with Abigail Guinan, Ethan Rutledge, JT Larkin

### Intermediate Logical Subsetting II: Qs 1-2

```{r}
n = 12345
vec_1 = sample(12, n, replace = TRUE)
head(vec_1)

vec_2 <- vec_1 == 3
#vec_1[vec_2]

```


#### Q1 
vec_2 <- vec_1 == 3

#### Q2
A visual inspection would be a bad idea because it would be very time consuming and likely that human error will effect the result

### Intermediate Logical Subsetting II: Qs 3-5

```{r}
length(vec_1)

sum(vec_1 == 3)

n = 10
vec_1 = sample(12, n, replace = TRUE)
paste0("Sum of elements with value 3: ", sum(vec_1 == 3))

```
#### Q3 
The count was different each time becuase the sample() argument sampled 10 numbers at random in the entire vector. Each time the sampling function is executed a new set of numbers are sampled. The likelihood that the number '3' within a random sample of 10 numbers from a vector with integers 1:12 and a length of >12,000 numbers, would definitely be variable. 

#### Q4
A logical test is the safest way to select entries with a value of 3 becuase it provides a good random sampling of the whole spread of the data. 

#### Q5
Logical subsetting by hand is a bad practice because it can result in biased results and likely to have an error.


### Basic Loops: Q6
```{r}
for (i in 1:10)
{
  print(paste0("This is loop iteration:",i))
}

```

#### Q6 
Code for modified loop. 
for (i in 1:10)
{
  print(paste0("This is loop iteration:",i))
}

### Intermediate Loops: Q7 

```{r}

n=6

for (i in 1:n)
{
  print(i)
}


```

#### Q7
n=6

for (i in 1:n)
{
  print(i)
}

### Intermediate Loops 2: Q8

```{r}

n=17

vec_1 = sample(10, n, replace = TRUE)
vec_1

#this is the correct code
for (i in 1:n)
{
  print(paste0("The element of vec_1 at index ", i, " is : ", vec_1[i]))
}

```

#### Q8 

for (i in 1:n)
{
  print(paste0("The element of vec_1 at index ", i, " is : ", vec_1[i]))
}

### Functions Q9

```{r}

create_and_print_vec = function(n, min = 1 , max = 10)
  { 
  vec_1 = sample(min:max, n, replace = TRUE)
  
  for (i in 1:n)
      { 
    print(
      paste0("The element at index ", i, " is : ", vec_1[i]))
      }
}

create_and_print_vec(10, min = 100, max = 200)

```

#### Q9 
create_and_print_vec = function(n, min = 1 , max = 10)
  { 
  vec_1 = sample(min:max, n, replace = TRUE)
  
  for (i in 1:n)
      { 
    print(
      paste0("The element at index ", i, " is : ", vec_1[i]))
      }
}