---
title: "Lab 07"
author: "Sonja"
date: "20/10/2021"
output: html_document
---

### Learning objectives, technical skills and concepts:

- Bootstrap
- Apply functions
- Building custom functions
- Resampling


Worked with:

Aliza, Abby, Justin


```{r setup, include=FALSE}
require(palmerpenguins)
head(penguins)
require(boot)
require(here)
```

```{r, subset Gentoo, echo=FALSE}

dat_gen = subset(penguins, species == "Gentoo")

na.x<-is.na(dat_gen$bill_length_mm)
          dg<- dat_gen$bill_length_mm[!na.x]

#or
dat_gen2<- na.omit(dat_gen$bill_length_mm)


```


Calculate a parametric 95% CI for mean bill length (in mm) for the Gentoo penguins in the penguins dataset from package palmerpenguins using your SSE function. For this calculation you should use Student’s t-distribution to calculate the critical values.
Q1-Q5


#### Q1
What is the sample size, n? Show the code you used for the calculation and remember to check for missing data.
```{r}
n = length(dg)
n
```

```{r, NA omit}
na.x<-is.na(dat_gen$bill_length_mm)
          dg<- dat_gen$bill_length_mm[!na.x]
          
```


#### Q2
What is the sample standard deviation? Show the code you used for the calculation.
```{r}
sd.gen=sd(dg)
sd.gen

```
#### Q3
What are the critical t-values? Show the R code you used for the calculation.

```{r}
alpha=0.05

tcrit=qt(alpha/2, 122, lower.tail=FALSE)
tcrit
-tcrit
```
#### Q4
What is the sample standard error? Show the R code you used for the calculation.

```{r}
sse_mean = function(x)
    { na.x<-is.na(x)
          x2<- x[!na.x]
                 sd.x2=sd(x2)
                  n=length(x2)
        sd.x2/(sqrt(n))
}

ssedg=sse_mean(dg)
ssedg
```


#### Q5
Finally, construct the CI and show the R code you used for the calculation.

Multiply the sample standard error by the critical t-value. This is the radius of the CI.
Express the CI as mean±radius

```{r}
#mean 
mean.x=mean(dg)

#radius
radius=tcrit*ssedg

#CI upper & lower
mean.x+radius
mean.x-radius

```
$47.5\pm 0.5501$


Calculate a bootstrap 95% CI for mean bill length (in mm) for the Gentoo penguins in penguins dataset from package palmerpenguins.
Use the boot() function from package boot()
```{r}
boot_mean = function(x, i)
{
  return(mean(x[i], na.rm = TRUE))
}

```


```{r}
myboot = 
  boot(
    data = dg,
    statistic = boot_mean,
    R = 10000)
print(myboot)
```


#### Q6-Q8

#### Q6

What is the CI? 
   2.5%    97.5% 
46.97071 48.05287 

#### Q7

Show the r code you used to call the boot() function.

```{r}
require(boot)
```


#### Q8 
Show the r code you used to calculate the upper and lower 2.5% quantiles.

```{r}
quantile(
  myboot$t,
  c(0.025, 0.975))
```
#### Q9
Show your completed rarefaction_sampler() function.

```{r, echo=FALSE}
moths = read.csv(here("data", "moths.csv"))

```



```{r}

rarefaction_sampler = function(input_dat, n_iterations)
{
 n_input_rows = nrow(input_dat)

  results_out = matrix(
    nrow = n_iterations,
    ncol = n_input_rows)

  # The outer loop: runs once for each bootstrap iteration.  index variable is i
  for(i in 1:n_iterations)
  {
    # The inner loop: simulates increasing sampling intensity
    # Sampling intensity ranges from 1 site to the complete count of
    # sites in the input data (n)
    for(j in 1: n_input_rows)
    {
      # sample the input data row indices, with replacement
      rows_j = sample(n_input_rows, size = j, replace=TRUE)

      # Creates a new data matrix
      t1 = input_dat[rows_j, ]

      # Calculates the column sums
      t2 = apply(t1, 2, sum)

      # Counts the number of columns in which any moths were observed
      results_out[i, j] = sum(t2 > 0)
    }
  }
  return(results_out)
}

rarefact = rarefaction_sampler(moths, 100)
head(rarefact)

```

#### Q10: 
What did you find most difficult about building the function?
I do think that I actually built the formula, but I had to edit parts of the formula. I think it was challenging connecting the function loop variables to the other loops. I really did not understand what the point of the challenge was. 


#### Q11
Show the code you used to perform the simulations and construct the curve

```{r}
rarefact = rarefaction_sampler(moths[,-1], 10000)
```

```{r}
rare_mean = apply(rarefact, 2, mean)
rare_quant = apply(rarefact, 2, quantile, probs=c(0.025, 0.975))
rare = t(rbind(rare_mean, rare_quant))
```


#### Q12
Include your rarefaction curve plot in your report. Show the R-code you used to create your plot.
```{r}
matplot(
  rare,
  type='l',
  lwd=3,
  xlab='# of sampling plots',
  ylab='moth species richness',
  main='Rarefaction Curve for Collecting Moths', col=c("blue","orange","purple"))

legend(
  'bottomright',
  legend=c('mean','2.5%','97.5%'),
  lty=c(1,2,3),  lwd=3, col=c("blue","orange","purple"), inset=c(.05,.05), title="Mean Curve & CIs",text.font=2)

```

#### Q13
About how many sites should you visit if you want to see all of the moth species? Explain your reasoning using your rarefaction curve figure.

Around 10-15 sites should capture all of the the moth species. Around 10-15 sites is where both quantile curves and the mean curve reach a total of 10 moths. But, by just looking at the bottom 2.5% quantile, perhaps at 20 sites there would be more certainty of capturing all moths.    
