---
title: "Pre-Lab 07"
author: "Sonja"
date: "20/10/2021"
output: html_document
---

### Learning objectives, technical skills and concepts:

- Bootstrap
- Apply functions
- Building custom functions
- Resampling

#### The apply() family of functions


apply() works on data frames or other 2-dimensional arrays of data, such as matrices.

It applies a function to either the rows or columns of the input data.

##### The important arguments are:

- X the 2-dimensional data, usually a data frame or a matrix
- MARGIN whether to apply the function to rows (MARGIN = 1) or columns (MARGIN = 2)
- FUN The function to apply to the rows or columns

```{r, Create simulated data}
dat = matrix(1:49, nrow = 7, ncol = 7)
print(dat)
```


```{r}
apply(dat, MARGIN = 1, FUN = min)
apply(dat, MARGIN = 1, FUN = max)
apply(dat, MARGIN = 2, FUN = mean)
```
More advanced usage
https://www.datacamp.com/community/tutorials/r-tutorial-apply-family



##### Data Files

We’ll work with a dataset containing standardized abundance estimates of 10 rare moth species across 24 sample plots in the pine barrens of southeast Massachusetts.

```{r}
require(here)
moths = read.csv(here("data", "moths.csv"))
head(moths)
```
#### The Bootstrap
In a statistical context, bootstrap refers to a resampling technique by with which we use sampling with replacement on a single sample of observations to simulate the process of taking many samples from a population.

Bootstrap and Monte-Carlo resampling
These two resampling techniques can help us learn something about the null and alternative hypotheses.

Bootstrapping creates a kind of alternative distribution, while Monte-Carlo resampling can characterize a null distribution.

- *The alternative distribution* corresponds to a range of possibilities given associations in the data we actually observed. *Sampling with replacement allows some values to appear more than once, and other samples to be left out.*

- *The null distribution* tells us something about what we could expect to see if there were no associations.

###### *One way to think of the difference is that*:

- Monte-Carlo resampling shuffles each column separately, creating new combinations of attribute values that were not observed in the original data.

- Bootstrapping samples an entire row; it does not create new data.


One of the most important uses of the bootstrap is calculating non-parametric confidence intervals for parameter estimates.

In this context, the bootstrap simulates the frequentist ideal of obtaining estimates from repeated similar experiments.


##### Moth Abundance: CIs

Let’s say we want to obtain a 95% confidence interval for the mean standardized abundance.

```{r}
hist(moths$anst)
```


1. Calculate the sample standard error.
     - You’ll need the sample standard deviation and the sample size to do this calculation.
2. Choose your significance level (usually 95%).
     - The value of alpha is just 1 - significance level.
3. Using your chosen value of alpha, calculate the critical t-values.
4. Multiply the sample standard error by the critical t-value. This is the radius of the CI.
5. Express the CI as mean±radius

```{r}
sse_mean = function(x)
    { na.x<-is.na(x)
          x2<- x[!na.x]
                 sd.x2=sd(x2)
                  n=length(x2)
        sd.x2/(sqrt(n))
      }

sse_mean(moths$anst)
length(moths$anst)
# alpha = 0.05


```
```{r}
alpha=0.05
qt(1-alpha/2, 23, lower.tail=FALSE)

  sse_mean = {na.x<-is.na(x)
          x2<- x[!na.x]
                 sd.x2=sd(x2)
                  n=length(x2)
        sd.x2/(sqrt(n))}

   mean.x<-mean(moths$anst, 23)
        tcrit=qt(1-0.05/2, 23, lower.tail=FALSE)  
        mean.x+tcrit
        mean.x-tcrit
  
ci_1 = function(x,alpha,n)
    { mean.x<-mean(x)
        tcrit=qt(1-alpha/2, n, lower.tail=FALSE)  
        mean.x+tcrit
        mean.x-tcrit
}

ci_1(moths$anst,0.05, 23)

```

#### A Simple Bootstrap Confidence Interval

An alternative to the parametric approach above is to use bootstrapping to obtain a confidence interval for the mean.

```{r}
m = 10000

# numeric() creates an vector of length m with all values initialized to zero
result = numeric(m)
head(result)
```

Perform the bootstrap

```{r}

for(i in 1:m)
{
  result[i] = mean(sample(moths$anst, replace=TRUE))
}


```

##### Calculate the quantiles

The vector result now contains 10,000 bootstrap sample means. We can calculate the mean of the bootstrap means and, more importantly, the 2.5% and 97.5% quantiles of the bootstrap distribution, as follows:

```{r}
mean(result)
quantile(result,c(0.025,0.975))
```

##### Bootstrap Interval Using boot()

```{r}
install.packages("boot")
```

```{r}
require(boot)
boot(data, statistic, R)
```
R is the number of resamplings you want to perform
data is the data object you want to resample. It can be a vector, matrix, or data.frame.
statistic is a function that returns the statistic of interest. You don’t put quotes around the name of the function.

We’ll usually need to write a custom function to meet the restrictions.
Be sure to check out the help entry. It’s a difficult read, so hopefully an example will illustrate the process:

##### Custom Mean Function
We want to calculate to mean, but we know that mean() can sometimes have trouble with NA values.

To use it with boot() we have to create a modified version of the mean() function that automatically excludes NAs:

```{r}
boot_mean = function(x, i)
{
  return(mean(x[i], na.rm = TRUE))
}
```


```{r}
myboot = 
  boot(
    data = moths$anst,
    statistic = boot_mean,
    R = 10000)
print(myboot)

str(myboot)
```

The output is interpreted as follows:

original is the original mean of the whole sample: mean(moths$anst).
bias is the difference between the original mean and the mean of the bootstrapped samples.
std.error is the standard deviation of the simulated values.

```{r}
mean(moths$anst)
myboot$t0
mean(myboot$t) - myboot$t0
sd(myboot$t)
```

```{r}
quantile(
  myboot$t,
  c(0.025, 0.975))
```
##### Rarefaction Curve
The principle behind rarefaction is that the number of species detected is related to sampling intensity:

The greater the sampling intensity (e.g., number of sample plots or sampling area) the greater the species richness.

```{r}
moth_dat = moths[,-1]
head(moth_dat)
```


```{r}
n = nrow(moth_dat) #number of rows or sample observations
m = 100 #number of bootstrap iterations
moth_result = matrix(
  nrow = m,
  ncol = n)
```


```{r}
n = nrow(moth_dat) #number of rows or sample observations

m = 100 #number of bootstrap iterations

moth_result = matrix(
  nrow = m,
  ncol = n)


# The outer loop: runs once for each bootstrap iteration.  index variable is i
for(i in 1:m)
{
  # The inner loop: simulates increasing sampling intensity
  # Sampling intensity ranges from 1 site to the complete count of sites (24)
  # index variable is j
  for(j in 1:n)
  {
    # sample the input data row indices, with replacement
    rows_j = sample(n, size = j, replace=TRUE)

    # Creates a new data matrix from the resampled rows.
    t1 = moth_dat[rows_j, ]

    # Calculates the column sums of the new data matrix.
    t2 = apply(t1, 2, sum)

    # Counts the number of columns in which any moths were observed
    moth_result[i, j] = sum(t2 > 0)
  }
}

head(moth_result)
```


```{r}
rarefaction_sampler = function(input_dat, n_iterations)
{
  n = nrow(moth_dat) #number of rows or sample observations
  m = 100 #number of bootstrap iterations

  moth_result = matrix(
    nrow = m,
    ncol = n)

  # The outer loop: runs once for each bootstrap iteration.  index variable is i
  for(i in 1:m)
  {
    # The inner loop: simulates increasing sampling intensity
    # Sampling intensity ranges from 1 site to the complete count of sites (24)
    # index variable is j
    for(j in 1:n)
    {

      # sample the input data row indices, with replacement
      rows_j = sample(n, size = j, replace=TRUE)

      # Creates a new data matrix
      t1 = moth_dat[rows_j, ]

      # Calculates the column sums
      t2 = apply(t1, 2, sum)

      # Counts the number of columns in which any moths were observed
      moth_result[i, j] = sum(t2 > 0)
    }
  }

  return(moth_result)
}

rarefact = rarefaction_sampler(moth_dat, 100)
head(rarefact)
```
```{r}
rarefaction_sampler = function(input_dat, n_iterations)
{
  n_input_rows = nrow(input_dat)

  results_out = matrix(
    nrow = n_iterations,
    ncol = n_input_rows)

  # The outer loop: runs once for each bootstrap iteration.  index variable is i
  for(i in 1:n_iterations)
  {
    # The inner loop: simulates increasing sampling intensity
    # Sampling intensity ranges from 1 site to the complete count of
    # sites in the input data (n)
    # index variable is j
    for(j in 1:n)
    {
      # sample the input data row indices, with replacement
      rows_j = sample(n, size = j, replace=TRUE)

      # Creates a new data matrix
      t1 = input_dat[rows_j, ]

      # Calculates the column sums
      t2 = apply(t1, 2, sum)

      # Counts the number of columns in which any moths were observed
      results_out[i, j] = sum(t2 > 0)
    }
  }
  return(results_out)
}

rarefact = rarefaction_sampler(moth_dat, 100)
head(rarefact)
```

### Q1 edit function

```{r}
rm(list = ls())
moths = read.csv(here("data", "moths.csv"))
moth_dat = moths[,-1]

nrow(moth_dat)
matrix(
    nrow = 100,
    ncol =24)



rarefaction_sampler = function(input_dat, n_iterations)
{
  n_input_rows = nrow(input_dat)

  results_out = matrix(
    nrow = n_iterations,
    ncol = n_input_rows)

  # The outer loop: runs once for each bootstrap iteration.  index variable is i
  for(i in 1:n_iterations)
  {
    # The inner loop: simulates increasing sampling intensity
    # Sampling intensity ranges from 1 site to the complete count of
    # sites in the input data (n)
    for(j in 1:n)
    {
      # sample the input data row indices, with replacement
      rows_j = sample(n, size = j, replace=TRUE)

      # Creates a new data matrix
      t1 = input_dat[rows_j, ]

      # Calculates the column sums
      t2 = apply(t1, 2, sum)

      # Counts the number of columns in which any moths were observed
      results_out[i, j] = sum(t2 > 0)
    }
  }
  return(results_out)
}

rarefact = rarefaction_sampler(moth_dat, 100)
head(rarefact)

```

```{r}

```

