---
title: "Confidence Intervals"
author: "Sonja"
date: "28/10/2021"
output: html_document
---

## Critical Values

Critical values are used for construction of parametric CIs.

For a two-sided 95% CI, they are the x-values that correspond to the locations on the x-axis, beyond which 5% of the total probability density lies.

For example, a 95% CI constructed from a Standard Normal distribution has critical Z values of ±1.96:


Q1 90% Standard Normal Critical Values

```{r}
qnorm(c(0.05, 0.95))
```

Q2-4 95% T-distribution critical values


Q2
Consult the help entry for qt() and calculate the critical values for df = 10. Show the R-code you used to perform the calculation.

```{r}

qt(c(0.025, .975), df=10, lower.tail = FALSE) 

```

Q3 (2 pts.): How many degrees of freedom are required for the 0.025% lower critical value of a t-distribution to match the 0.025% lower critical z-value (from the standard normal) to within one decimal place? Show the R-code you used to perform the calculation.

```{r}

qt(c(0.025, .975), df=61
   , lower.tail = FALSE)

```

Q4 (1 pt.): How many degrees of freedom are required for the 0.025% lower critical value of a t-distribution to match the 0.025% lower critical z-value (from the standard normal) to within two decimal places? Show the R-code you used to perform the calculation.

```{r}

qt(c(0.025, .975), df=473
   , lower.tail = FALSE)

```

Q5 What are the critical t-values you would need to know to construct a 95% CI on the mean?

```{r}

qt(c(0.025, .975),
   df=49, 
   lower.tail = FALSE)


```

Q6: Construct the interval. Show the R-code you used to perform the calculation.

```{r}

mean=10
sd=3.14
n=50

tcrit=2.009575
sse=sd/(sqrt(n))

#radius
radius=tcrit*sse

#CI upper & lower
mean+radius
mean-radius
```

