---
title: "in class prob calc"
author: "Sonja"
date: "14/10/2021"
output: html_document
---

In-Class Probability Calculations and Confidence Intervals

Objectives
Practice calculating probabilities using R’s distribution functions.

R has built-in functions for calculating the probabilities of events for many parametric distributions.

The d-functions (dbinom, dnorm, dpois, etc.) calculate the probability density (or mass) of a single event. For a discrete distribution like the binomial, the probability mass of an event is the probability that the event happens.

Each of the d-functions has different arguments, depending on the specific parameters of the distribution in question.

For example, recall the Poisson distribution, which has 1 parameter: lambda.

Using dpois I can ask what is the probability that I observe a count of exactly 7 if I have a poisson-distributed population with lambda = 10.4:

```{r}
dpois(x = 7, lambda = 10.4)

```
Recall the binomial distributions, and their two parameters:

n: the number of trials. R calls this size.
p: the probability of success on each trial. R calls this prob.
For example, if I had a set of six forest plots, and I knew from earlier observations that when I do a survey, I observe birds in about 2/3 of the plots (i observe 4 presences out of 6 sites), I could use a binomial distribution to model my set of plots.

Q1= n=6, prob=2/3

Q2
```{r}
#Q2
dbinom(x=4,6,2/3)
#Q3
dbinom(x=0,6,2/3)

```

Cumulative probability: The p-functions
The d-functions calculate the probability density or mass of observing a specific event.

I can use the p-functions to calculate the probabilities of observing a range of events.

For example, I can use ppois to ask what is the probability that I observe a count of 7 or fewer if I have a poisson-distributed population with lambda = 10.4:

```{r}
ppois(q = 7, lambda = 10.4)
```
Law of Total Probability and Complementary Events
I can also use it to ask what the probability of observing a count greater than 7 using the law of total probability:


```{r}
1 - ppois(q = 7, lambda = 10.4)
```



```{r}
1 - ppois(q = 6, lambda = 10.4)
```

```{r}
pbinom(4,6,2/3)
```
```{r}
1-pbinom(3,6,2/3)
```

