---
title: 'Models 1: t-test'
author: "Sonja"
date: "2/11/2021"
output: html_document
---

```{r}
require(palmerpenguins)

```

```{r}
dat_ade = droplevels(subset(penguins, species == "Adelie"))

```

```{r}
hist(dat_ade$body_mass_g, main = "Adelie Penguins: Body Mass", xlab = "body mass (g)")

```

```{r}
boxplot(body_mass_g~sex, data = dat_ade, main="Body Mass of Adelie Penguins", ylab="body mass (g)")
```

 Recall that by default, when you provide the t.test() function with a group of measurements, it performs a one-sample t-test of the null hypothesis that the mean of the the group is equal to zero.
 
 
 Perform a one-sample t-test of the alternative hypothesis that female Adelie penguins have a body mass different from zero grams. Note that this is a very silly alternative hypothesis. Is this a one- or two-tailed test? Show your R-code.
 
```{r}
pen_fem = droplevels(subset(dat_ade, sex != "male"))

t.test(pen_fem$body_mass_g, y = NULL,
       alternative = c("greater"),
       mu = 0,
       conf.level = 0.95)

```
 
 Q3 Describe your conclusions based on the p-value of the t-test.
 the pvalue is very very small, so we could say that we can reject the null hypothesis that the Adelie female penguins weigh more than 0 g. Based on our confidence interval of 95% we find that if the true body mass was 0 grams it would be very unlikely that we would find the value we found from our sample of female Adelie penguins. 
 
 Q4
 Now, conduct a slightly less silly test: perform a one-sample t-test of the null hypothesis that male Adelie penguins have a mean body mass greater than 4000 grams. Is this a one- or two-tailed test?

 
```{r}
pen_male = droplevels(subset(dat_ade, sex != "female"))

t.test(pen_male$body_mass_g, y = NULL,
       alternative = "greater",
       mu = 4000, 
       conf.level = 0.95)

```
 
 Q5 our p-value is 1 meaning that we cannot reject the null hypothesis that the true mean of the Adelie female penguins is greater than 4000g
 
```{r}
t.test(body_mass_g~sex, data=dat_ade,
       alternative = "two.sided")
```


