---
title: "Using Models 1"
author: "Sonja"
date: "3/11/2021"
output: html_document
---

## Learning Objectives

Conduct commonly used tests on categorical data


### Key Concepts

- Analysis of categorical data: 1- and 2-sample tests

- Testing assumptions: normality

- Interpreting test output

```{r, echo=FALSE}
require(here)
catrate = read.csv(here("data", "catrate.csv"))

```

### Numerical and Graphical exploration

##### Q1

The values of the cat.rate collumn does not appear to be normally distributed. 

```{r}
hist(catrate$cat.rate, main = "Histogram of Catastophe Rates", xlab="Catastophe Rate")
```


### Check for Normality

##### Q2

p-value = 0.04097

```{r}
shapiro.test(catrate$cat.rate)
```
##### Q3 

The null hypothesis for the shapiro test is that the data were sampled from a normally-distributed population. 


##### Q4 

Based on the Shapiro test there is strong evidence that the sample came from a non-normally distributed population. 


### One-Sample Tests: Tests for Difference From Expectation

#### Parametric One-Sample Test: The t-test

##### Q5

```{r}

t.test(catrate$cat.rate,
       alternative = c("two.sided"),
       mu = 0.2857143,
       conf.level = 0.95)

```


##### Q6 

The null hypothesis is that the true mean is equal to 0.2857143. 

##### Q7 

This is a two-tailed test. 

##### Q8

If we were to go out and sample for the catastrophic rate of salamander eggs and we happened to observed many dry ponds and therefore a recorded a high catastrophic rate, it is possible that we what we observed is not true for the entire system and that in actuality the catastrophic rate is smaller. 

##### Q9 

The confidence interval for the difference between the null hypothesis and alternative hypothesis means is 0.3526250 - 0.7261295 and it does not include 0. 

##### Q10 

Based on the results from the t-test there is strong evidence to reject the null hypothesis. The p-value is 0.0119 meaning that it is very unlikely to observe the catastrophic rate we observed if the mean of the catastrophic rate was actually 0.28


#### One-sided Alternative Hypothesis


##### "greater"


```{r}
t.test(catrate$cat.rate, y = NULL,
       alternative = "greater",
       mu = 0.2857143,
       conf.level = 0.95)
```
##### "less"

```{r}
t.test(catrate$cat.rate, y = NULL,
       alternative = "less",
       mu = 0.2857143,
       conf.level = 0.95)
```

It shows that it is very unlikely that the carastrophic rate would be less than .28. 

### Non-Parametric One-Sample Test: The Wilcoxon Rank Sum Test

##### Q11

```{r}
wilcox.test(catrate$cat.rate, 
            mu = 2 / 7)

```
##### Q12

The p value from the Wilcox test is 0.0063, which is lower that the t-test's p value 


##### Q13

Yes there is strong evidence to reject the null hypothesis because the low p-value of 0.0063 from the Wilcox test means that the true mean is not equal to 2/7.


##### Q14 

The Wilcox test and the one-sample t tests show that it is very unlikely that the mean of the catastrophic reproduction rate is 2/7. The sample mean is actually 0.5393 which suggests the the catastrophic rate is not closely related to the rate of lake filling. 

##### Q15 

The Wilcox test is definitely better for this data set because of the clear violations of the normality assumption that is required for a t-test. 

##### Q16 


### Comparing two sample means


```{r, echo=FALSE}
require(palmerpenguins)
penguin_dat = droplevels(subset(penguins, species != "Gentoo"))
```
##### Q16


```{r}
# Extract the Adelie penguin data
dat_adelie = subset(penguin_dat, species == "Adelie")
# Extract the Chinstrap penguin data
dat_chin = subset(penguin_dat, species == "Chinstrap")
```

```{r}

sh<-aggregate(flipper_length_mm ~ species,
          data=penguin_dat, 
          FUN=function(x) shapiro.test(x)$p.value)

sh

```

##### Q17 

For Adelie penguins the Shapiro test p value was 0.72 and for the Chinstrap was p value 0.81 meaning that the null hypothesis cannot be rejected and that the data for both groups is normally distributed.


##### Q18 

```{r, echo=FALSE}
par(mfrow = c(1, 2))

hist(dat_adelie$flipper_length_mm, main="Adelie penguin\n flipper length", xlab = "Flipper length (mm)", col = "goldenrod3")

hist(dat_chin$flipper_length_mm, main="Chinstrap penguin\n flipper length", xlab = "Flipper length (mm)", col = "darkorchid3")
```
##### Q19

The alternative hypothesis is that the mean flipper length of Adelie penguins and Chinstrap penguins is different. 

##### Q20 

```{r}

t.test(flipper_length_mm~species, data=penguin_dat,
       alternative = "two.sided")

```

